### 2024-01-05 R FIle.qmd
adding files to repo

---
title: "testibng"
format: html
editor: visual
---

### Loading Packages

```{r}
library(tidytuesdayR)
library(tidyverse)
```

### Gathering Data

```{r}
# Union Data
tuesdata <- tidytuesdayR::tt_load(2023, week = 36)

demographics <- tuesdata$demographics
wages <- tuesdata$wages
states <- tuesdata$states

# Election Data
tuesdata <- tidytuesdayR::tt_load(2023, week = 45)

house <- tuesdata$house
```

### Cleaning Data

First, lets get an idea of what our data looks like.

```{r}
head(demographics)
head(wages)
head(states)
head(house)
```

From the head() function we can already see some tidying that needs to be done, mostly in the `house` table.

#### Items that need to be done

-   Sentence case

### EDA

```{r}
states |> 
  ggplot(aes(x = year, y = p_covered, color = state)) +
  geom_point()
```

```{r}
house <- house |>
  mutate(state = str_to_sentence(state))
```

```{r}
states |> 
  filter(sector == "Total") |> 
  left_join(house, by = c("year", "state")) |> 
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
  filter(year > 2012) |> 
  ggplot(aes(x = party, y = p_members)) +
  geom_boxplot()
```

```{r}
test <- states |> 
  filter(sector == "Total") |> 
  left_join(house, by = c("year", "state")) |> 
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
  filter(year > 2012)
```

```{r}
t.test(p_members ~ party, data = test, conf.level = .9)
```