### 2024-01-10 Code/Presentation.qmd
created final presentation

---
title: "WIDS 2024 Final Presentation"
author: "Michael Taffe, Melanie Klein, Ramiro Arevalo"
format: 
  revealjs:
    theme: simple
editor: visual
---

## Project Abstract

## About the Data

* **Union data**
  * Gathered from Tidy Tuesday, although the original 
* **House election data**
  *
* **Wage**
  *

## EDA

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Models

## Conclusion

## Where we would take our project in the future